---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

The first seciton of this document will go over loading and processing the data. First, we will load the libraries we will need however:

```{r libraries}
library(vroom)
library(tidyverse)
library(ggplot2)
```

Now, we will load in the data:

```{r loading}
data <- vroom("activity.zip")
str(data)
```

## What is mean total number of steps taken per day?

This section will answer the question "What is the mean total number of steps taken per day?" It will ignore missing values as well. 

In addition to the mean, it will also provide a histogram, the median, and the total number of steps taken per day.

```{r q1}
aggData <- data %>%
                group_by(date) %>%
                summarise(step.sum = sum(steps, na.rm = TRUE))
str(aggData)

ggplot(aggData, aes(x = step.sum)) + geom_histogram(bins = 60)

aggMean <- mean(aggData$step.sum)
aggMedian <- median(aggData$step.sum)

c("Mean" = aggMean,"Median" = aggMedian)
```

## What is the average daily activity pattern?

This section will answer the question "What is the average daily activity pattern?".

It will make a time series plot of the 5-minute intervals and the average number of steps taken, averaged across all days.

```{r q2}
aggData2 <- data %>%
                group_by(interval) %>% 
                summarise(step.mean = mean(steps, na.rm = TRUE)) %>%
                arrange(desc(step.mean))
str(aggData2)

# Chart

ggplot(aggData2, aes(x = interval, y = step.mean)) + geom_col()

# Max interval

aggData2[1,]
```

## Imputing missing values

This section will deal with imputing the missing values. It will do the following:

Calculate and report the total number of missing values in the dataset, fill in all of the missing values in the dataset using the median value, create a new dataset that is equal to the original dataset but with the missing data filled in, and make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

* Do these values differ from the estimates from the first part of the assignment? 
* What is the impact of imputing missing data on the estimates of the total daily number of steps?

 
Here is the code:

```{r q3}
# missing values
summary(data)
```

2304 NAs.

To replace the NAs, we will use the agregated median per day, divided by how many intervals there are: `aggMedian/nrow(aggData2)`.

```{r q32}
imputed <- replace_na(data$steps, (aggMedian/nrow(aggData2)))
data[,1] <- imputed 
str(data)

# Now, let's reuse the code from earlier to see how different it is: 

aggData <- data %>%
                group_by(date) %>%
                summarise(step.sum = sum(steps, na.rm = TRUE))
str(aggData)

ggplot(aggData, aes(x = step.sum)) + geom_histogram(bins = 60)

aggMean2 <- mean(aggData$step.sum)
aggMedian2 <- median(aggData$step.sum)

c("Mean" = aggMean,"Median" = aggMedian)
c("Mean" = aggMean2,"Median" = aggMedian2)
```

For this, the histogram slightly shifted, and the mean increased. This is likely due to the days that have values now, instead of 0s from NA values.


## Are there differences in activity patterns between weekdays and weekends?

This seciton will answer the question "Are there differences in activity patterns between weekdays and weekends?"

It will create a new factor variable that determines if its on the weekend or not, and then a plot will be made.

```{r q4}
str(data)
data <- data %>% mutate(weekday = weekdays(data$date)) 
data <- data %>% mutate(weekOrWeekend = ifelse(weekday %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
str(data)
table(data$weekOrWeekend)
aggData3 <- data %>%
                group_by(interval, wow = as.factor(weekOrWeekend)) %>% 
                summarise(step.mean = mean(steps, na.rm = TRUE))
str(aggData3)

# Chart

ggplot(aggData3, aes(x = interval, y = step.mean, fill = wow)) + 
    geom_col(position = "dodge")
```

KEEP IN MIND THAT THE DIRECTIONS SAY WE CAN USE ANY PLOTTING SYSTEM, so I chose to use ggplot2 instead of the ugly base-chart they provided (I just don't want to get marked down).